---
title: "HW2"
author: "Logan"
date: "9/9/2020"
output:
  word_document: default
  html_document: default
---

```{r hw2, include=FALSE}
library("Sleuth3")
library("tidyr")
library("tidyverse")
```

## 2-3

False. As the sample size increases, the sampling distribution approaches a normal shape, regardless of the population distribution due to the the Central Limit Theorem. However, this does not mean that the histogram of the sample is necessarily normal.

## 2-6

Y-bar is the sample statistic that will be used to approximate the value of mu, which is the parameter that we are trying to estimate. A hypothesis must be about the unknown parameter.

## 2-7

The hypothesis that the mean difference is zero the null hypothesis. We can say with 95% confidence that the difference in head circumference is between 0.61 and 1.19 cm. Since 0 cm is not within that 95% confidence interval, we have reason to believe that the null hypothesis is not true. We can say that the p-value is below 0.05, which refers to the 5% chance that the statistic we got was strictly due to chance.

## 2-13

``` {r 2-13, echo = FALSE, message = FALSE}
#install.packages("ANOVAreplication")
library("ANOVAreplication")
data13 <- (ex0112)
fishoil <- filter(data13, Diet == "FishOil")
regularoil <- filter(data13, Diet == "RegularOil")
meanfishoil <- mean(fishoil[1:7, 1:1])
meanregularoil <- mean(regularoil[1:7, 1:1])
sdfishoil <- sd(fishoil$BP)
sdregularoil <- sd(regularoil$BP)

pooledsd <- pooled.sd(data13)

standardE <- pooledsd*sqrt((1/7)+(1/7))

t95 <- 2.179
halfwidth <- t95 * standardE
lowerlimit <- (meanfishoil - meanregularoil) - halfwidth
upperlimit <- (meanfishoil - meanregularoil) + halfwidth

tstat <- ((meanfishoil - meanregularoil) - 0) / standardE

#t.test(fishoil$BP,regularoil$BP, var.equal = T)

```
a.
The average of the fish oil group is __`r meanfishoil`__.  
The average of the regular oil group is __`r meanregularoil`__.  
The sample standard deviation of the fish oil group is __`r sdfishoil`__.  
The sample standard deviation of the regular oil group is __`r sdregularoil`__.

b.
The pooled estimate of the standard deviation is __`r pooledsd`__.

c.
The standard error is __`r standardE`__.

d.
The degrees of freedom associated with the pooled estimate of the standard deviation is equal to $$n_2+n_1-2$$, which in this case is equal to $$7+7-2=12$$. The 97.5th percentile of the t-distribution with 12 degrees of freedom is __2.179__.

e.
A 95% confidence interval for the difference in population means has a lower limit of __`r lowerlimit`__ and an upper limit of __`r upperlimit`__.

f.
The *t*-statistic for testing equality is __`r tstat`__.

e.
Based off of the Student's T-Table and a *t*-statistic of `r tstat`, the one-sided *p*-value is __0.005__.

These results are confirmed by a t-test, which gives a two-sided p-value:

``` {r ttest, echo = FALSE}
t.test(fishoil$BP,regularoil$BP, var.equal = T)
```


## 2-18


a. 
```{r 2-18a, echo = FALSE}
data18 <- (ex0218)
data1976 <- filter(data18, Year == "1976")
data1978 <- filter(data18, Year == "1978")
boxplot(data1976$Depth, data1978$Depth,
        main = "Beak Depths by Year",
        xlab = "1976                                                    1978",
        ylab = "Beak Depth (mm)")
```
b.
```{r 2-18b, echo = FALSE}
#Do a t-test with a one-sided p-value (omit alternative = "greater" for two-sided)
#do the 1978 data first because we use mu2 - mu1
t.test(data1978$Depth, data1976$Depth, var.equal = T, alternative = "greater")
```
As provided in the above *t*-test, the one-sided *p*-value is __1.617e-06__.

c.
```{r 2-18c, echo = FALSE}
#Do a t-test with a two-sided p-value (omit alternative = "greater" for two-sided)
#do the 1978 data first because we use mu2 - mu1
t.test(data1978$Depth, data1976$Depth, var.equal = T)
```
As provided in the above *t*-test, the two-sided *p*-value is __1.617e-06__.

d.
As provided by the *t*-tests given above, the mean of the 1978 data is 10.138202 mm and the mean of the 1976 data is 9.600799 mm, the difference of which is __0.537403 mm__. A 95% confidence interval for the amount by which the 1978 mean exceeds the 1976 mean is between __0.3123571__ and __0.7624495__.

e.
The populations are not independent of each other due to the fact that the finches which survived into 1978 were also part of the population studied in 1976.

## 2-22

```{r 2-22afqt, echo = TRUE}
data22 <- (ex0222)
male <- filter(data22, Gender == "male")
female <- filter(data22, Gender == "female")
boxplot(male$AFQT, female$AFQT,
        main = "AFQT Scores by Gender",
        xlab =  "Male                                                Female",
        ylab = "AFQT Score")

t.test(male$AFQT, female$AFQT, var.equal = T, alterative = "two.sided")
```
The AFQT scores show a moderate difference between males and females, with a male average score of __55.44625__ and a female average score of __53.40579__. The difference between these two sample means (male-female) is only __2.04046__ points. However, the two-sided *p*-value given by the *t*-test is 0.06715, which means that there is only moderate evidence of a difference. We do not reject the null hypothesis since our 95% confidence interval includes the 0 value of no assumed difference.

```{r 2-22arith, echo = TRUE}
boxplot(male$Arith, female$Arith,
        main = "Arithmetic Reasoning Scores by Gender",
        xlab =  "Male                                                Female",
        ylab = "Arithmetic Reasoning Score")

t.test(male$Arith, female$Arith, var.equal = T, alterative = "two.sided")
```
There appears to be a convincing difference between the arithmetic reasoning scores between men and women. The average of the male scores is __19.52297__ and the average of the female scores is __17.48592__. The difference between these sample means (male-female) is __2.03705__. The two-sided *p*-value is small enough to convince us that the null hypothesis is not correct. Furthermore, the 0 value of no assumed difference does not exist within the 95% confidence interval.

```{r 2-22word, echo = TRUE}
boxplot(male$Word, female$Word,
        main = "Word Knowledge Scores by Gender",
        xlab =  "Male                                                Female",
        ylab = "Word Knowledge Score")

t.test(male$Word, female$Word, var.equal = T, alterative = "two.sided")
```
There is no significant difference between the word knowledge scores between men and women. The average of the male scores is __26.54594__ and the average of the female scores is __26.56808__. The difference between these sample means (male-female) is __-0.02214__. The two-sided *p*-value is very large, which indicates to us that the null hypothesis is correct. Furthermore, the 0 value of no assumed difference exists almost exactly in the center of the 95% confidence interval.

```{r 2-22parag, echo = TRUE}
boxplot(male$Parag, female$Parag,
        main = "Paragraph Comprehension Scores by Gender",
        xlab =  "Male                                                Female",
        ylab = "Paraggraph Comprehension Score")

t.test(male$Parag, female$Parag, var.equal = T, alterative = "two.sided")
```
There appears to be a convincing difference between the paragraph comprehension scores between men and women. The average of the male scores is __10.92037__ and the average of the female scores is __11.48983__. The difference between these sample means (male-female) is __-0.56946__. The two-sided *p*-value is small enough to convince us that the null hypothesis is not correct. Furthermore, the 0 value of no assumed difference does not exist within the 95% confidence interval.

```{r 2-22math, echo = TRUE}
boxplot(male$Math, female$Math,
        main = "Mathematical Knowledge Scores by Gender",
        xlab =  "Male                                                Female",
        ylab = "Mathematical Knowledge Score")

t.test(male$Math, female$Math, var.equal = T, alterative = "two.sided")
```

There appears to be a convincing difference between the mathematical knowledge scores between men and women. The average of the male scores is __14.56738__ and the average of the female scores is __13.81534__. The difference between these sample means (male-female) is __0.75204__. The two-sided *p*-value is small enough to convince us that the null hypothesis is not correct. Furthermore, the 0 value of no assumed difference does not exist within the 95% confidence interval.