---
title: "HW5"
author: "Logan Bennett"
date: "10/7/2020"
output: html_document
---

```{r setup, include=FALSE}
library("Sleuth3")
library("tidyr")
library("tidyverse")
library("ANOVAreplication")
library("knitr")
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 5.4


## Problem 5.10


## Problem 5.16

```{r 16, echo = FALSE}
data16 <- (case0502)
#change dataframe into a linear model with values and groups > lm(y ~ group, data) (use aov or lm function)
spocks <- lm(Percent ~ Judge, data16)
anova(spocks)
```
See the above ANOVA table for the Spock Trial data.

## Problem 5.17

|Source|d.f.|Sum of Squares|Mean Square|F-statistic|p-vale|
|-----|------|------|----|------|-----|
|Between groups|7|35819|5117|3.5|0.009942|
|Within groups|24|35088|1462| | |
|Total|31|70907| | | |

From the above ANOVA table and the *p*-value of **0.009942** we have convincing evidence that at least one of the means of the **8** group means is different from the others.

## Problem 5.18

a.

```{r 18 a, echo = FALSE}
data18 <- (ex0518)
#filter data by treatment
c50 <- filter(data18, Treatment == "CPFA50")
c150 <- filter(data18, Treatment == "CPFA150")
c300 <- filter(data18, Treatment == "CPFA300")
c450 <- filter(data18, Treatment == "CPFA450")
c600 <- filter(data18, Treatment == "CPFA600")
control <- filter(data18, Treatment == "Control")
#Calculate the mean for each of the 6 groups
mean50 <- mean(c50$Protein)
mean150 <- mean(c150$Protein)
mean300 <- mean(c300$Protein)
mean450 <- mean(c450$Protein)
mean600 <- mean(c600$Protein)
meancontrol <- mean(control$Protein)
#calculate the grand mean and add a column for it
grandmean <- mean(data18$Protein)
equalmean <- c(rep(grandmean, 30))
data18 <- cbind(data18, equalmean)
#Add a column to data frame for equal means residual
equalres <- c(data18$Protein - grandmean)
data18 <- cbind(data18, equalres)
#Add columns for the separate means and separate means residual then add columns to data frame
sepmean <- c(rep(mean50, 3), rep(mean150, 3),rep(mean300, 3),rep(mean450, 3),rep(mean600, 3),rep(meancontrol, 15))
data18 <- cbind(data18, sepmean)
sepres <- c((c50$Protein - mean50),(c150$Protein - mean150),(c300$Protein - mean300),(c450$Protein - mean450),(c600$Protein - mean600),(control$Protein - meancontrol))
data18 <- cbind(data18, sepres)
#Plot residuals vs estimated means vs separate residuals
plot(data18$sepmean, data18$sepres,
     main = "Residual Plot of Estimated Mean vs. Residuals",
     xlab = "Treatment Mean",
     ylab = "Residual (x10)")
#Plot residuals vs day
plot(as.integer(data18$Day), data18$sepres,
     main = "Residual Plot of Day vs. Residuals",
     xlab = "Day",
     ylab = "Residual (x10)")
```
Based on Display 5.15 from the book, these residual plots demonstrate a non-constant variance. A logit transformation might help, but nothing indicates that the methods of this chapter are not appropriate.

b.

The estimated means for the model with 10 independent samples are given below in the table. Based on an ANOVA F-test, also summarized below, at least one of the 10 means is different from the others, due to the low *p*-value of **7.154e-05**.

```{r 18 b, echo = FALSE}
data18b <- (ex0518)
#Calculate means by TrtDayGroup and add a column to the data frame
tdg1 <- mean(filter(data18b, TrtDayGroup == "Group1")$Protein)
tdg2 <- mean(filter(data18b, TrtDayGroup == "Group2")$Protein)
tdg3 <- mean(filter(data18b, TrtDayGroup == "Group3")$Protein)
tdg4 <- mean(filter(data18b, TrtDayGroup == "Group4")$Protein)
tdg5 <- mean(filter(data18b, TrtDayGroup == "Group5")$Protein)
tdg6 <- mean(filter(data18b, TrtDayGroup == "Group6")$Protein)
tdg7 <- mean(filter(data18b, TrtDayGroup == "Group7")$Protein)
tdg8 <- mean(filter(data18b, TrtDayGroup == "Group8")$Protein)
tdg9 <- mean(filter(data18b, TrtDayGroup == "Group9")$Protein)
tdg10 <- mean(filter(data18b, TrtDayGroup == "Group10")$Protein)
#Create the column of means
Estimated_Means <- c(rep(tdg1, 3), rep(tdg2, 3), rep(tdg3, 3), rep(tdg4, 3), rep(tdg5, 3), rep(tdg6, 3), rep(tdg7, 3), rep(tdg8, 3), rep(tdg9, 3), rep(tdg10, 3))
#Add the column to the data frame
data18b <- cbind(data18b, Estimated_Means)
#Create a Table for the estimated means of the 10 independent samples
kable(data18b)
#Calculate the ANOVA F-test for a linear model of the data frame
tdglm <- lm(Protein ~ TrtDayGroup, data18b)
anova(tdglm)
```

c.




## Problem 5.23
