---
title: "HW5"
author: "Logan Bennett"
date: "10/7/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
library("Sleuth3")
library("tidyr")
library("tidyverse")
library("ANOVAreplication")
library("knitr")
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 5.4

When the sample sizes are different, a weighted average of the pooled estimate of the standard deviation is appropriate since estimates from larger samples will be given more weight. The pooled estimate will then be a better approximation if all the population variances are equal because it includes data from all the groups instead of isolating only two and ignoring the majority of the data.

## Problem 5.10

The role of the degrees of freedom of the *F*-distribution associated with the *F*-statistic is to gather a statistic from the correct distribution, as they fluctuate according to the sample sizes and number of groups being compared. Smaller degrees of freedom create more variability in the distribution, which can skew the F-ratio to a number further away from 1.

## Problem 5.16

```{r 16, echo = FALSE}
data16 <- (case0502)
#change dataframe into a linear model with values and groups > lm(y ~ group, data) (use aov or lm function)
spocks <- lm(Percent ~ Judge, data16)
anova(spocks)
```
See the above ANOVA table for the Spock Trial data.

## Problem 5.17

|Source|d.f.|Sum of Squares|Mean Square|F-statistic|p-vale|
|-----|------|------|----|------|-----|
|Between groups|7|35819|5117|3.5|0.009942|
|Within groups|24|35088|1462| | |
|Total|31|70907| | | |

From the above ANOVA table and the *p*-value of **0.009942** we have convincing evidence that at least one of the means of the **8** group means is different from the others.

## Problem 5.18

a.

```{r 18 a, echo = FALSE}
data18 <- (ex0518)
#filter data by treatment
c50 <- filter(data18, Treatment == "CPFA50")
c150 <- filter(data18, Treatment == "CPFA150")
c300 <- filter(data18, Treatment == "CPFA300")
c450 <- filter(data18, Treatment == "CPFA450")
c600 <- filter(data18, Treatment == "CPFA600")
control <- filter(data18, Treatment == "Control")
#Calculate the mean for each of the 6 groups
mean50 <- mean(c50$Protein)
mean150 <- mean(c150$Protein)
mean300 <- mean(c300$Protein)
mean450 <- mean(c450$Protein)
mean600 <- mean(c600$Protein)
meancontrol <- mean(control$Protein)
#calculate the grand mean and add a column for it
grandmean <- mean(data18$Protein)
equalmean <- c(rep(grandmean, 30))
data18 <- cbind(data18, equalmean)
#Add a column to data frame for equal means residual
equalres <- c(data18$Protein - grandmean)
data18 <- cbind(data18, equalres)
#Add columns for the separate means and separate means residual then add columns to data frame
sepmean <- c(rep(mean50, 3), rep(mean150, 3),rep(mean300, 3),rep(mean450, 3),rep(mean600, 3),rep(meancontrol, 15))
data18 <- cbind(data18, sepmean)
sepres <- c((c50$Protein - mean50),(c150$Protein - mean150),(c300$Protein - mean300),(c450$Protein - mean450),(c600$Protein - mean600),(control$Protein - meancontrol))
data18 <- cbind(data18, sepres)
#Plot residuals vs estimated means vs separate residuals
plot(data18$sepmean, data18$sepres,
     main = "Residual Plot of Estimated Mean vs. Residuals",
     xlab = "Treatment Mean",
     ylab = "Residual (x10)")
#Plot residuals vs day
plot(as.integer(data18$Day), data18$sepres,
     main = "Residual Plot of Day vs. Residuals",
     xlab = "Day",
     ylab = "Residual (x10)")
```
Based on Display 5.15 from the book, these residual plots demonstrate a non-constant variance. A logit transformation might help, but nothing indicates that the methods of this chapter are not appropriate.

b.

The estimated means for the model with 10 independent samples are given below in the table. Based on an ANOVA F-test, also summarized below, at least one of the 10 means is different from the others, due to the low *p*-value of **7.154e-05**.

```{r 18 b, echo = FALSE}
data18b <- (ex0518)
#Calculate means by TrtDayGroup and add a column to the data frame
tdg1 <- mean(filter(data18b, TrtDayGroup == "Group1")$Protein)
tdg2 <- mean(filter(data18b, TrtDayGroup == "Group2")$Protein)
tdg3 <- mean(filter(data18b, TrtDayGroup == "Group3")$Protein)
tdg4 <- mean(filter(data18b, TrtDayGroup == "Group4")$Protein)
tdg5 <- mean(filter(data18b, TrtDayGroup == "Group5")$Protein)
tdg6 <- mean(filter(data18b, TrtDayGroup == "Group6")$Protein)
tdg7 <- mean(filter(data18b, TrtDayGroup == "Group7")$Protein)
tdg8 <- mean(filter(data18b, TrtDayGroup == "Group8")$Protein)
tdg9 <- mean(filter(data18b, TrtDayGroup == "Group9")$Protein)
tdg10 <- mean(filter(data18b, TrtDayGroup == "Group10")$Protein)
#Create the column of means
Estimated_Means <- c(rep(tdg1, 3), rep(tdg2, 3), rep(tdg3, 3), rep(tdg4, 3), rep(tdg5, 3), rep(tdg6, 3), rep(tdg7, 3), rep(tdg8, 3), rep(tdg9, 3), rep(tdg10, 3))
#Add the column to the data frame
data18b <- cbind(data18b, Estimated_Means)
#Create a Table for the estimated means of the 10 independent samples
kable(data18b)
#Calculate the ANOVA F-test for a linear model of the data frame
tdglm <- lm(Protein ~ TrtDayGroup, data18b)
anova(tdglm)
```

c.

A comparison of the 10-mean model and the 6-mean model is given below.

```{r 18 c, echo = TRUE}
#Perform F-tests for the data by Treatment (6 means) and by TrtDayGroup (10 means)
meanmodel6 <- lm(Protein ~ Treatment, data18b)
anova(meanmodel6)
meanmodel10 <- lm(Protein ~ TrtDayGroup, data18b)
anova(meanmodel10)
```
ANOVA tables from tow F-tests are shown above, the first performed for the 6-mean model and the second for the 10-mean model. The means from the control group on different days do differ from one another because there is a much smaller *p*-value from the F-test run for 10 separate means (**p = 7.154e-05**) than from the F-test run for 6 separate means (**p = 0.003196**).


## Problem 5.23

The statistical procedures performed to determine whether or not the Tyrannosaurus Rex was warm-blooded or not is shown below. Visualization of the oxygen levels of each respective bone will be shown, along with results from an F-test performed on the data.

```{r 23 visualize, echo = FALSE}
data23 <- (ex0523)
#Compare the means visually
bone1 <- filter(data23, Bone == "Bone1")
bone2 <- filter(data23, Bone == "Bone2")
bone3 <- filter(data23, Bone == "Bone3")
bone4 <- filter(data23, Bone == "Bone4")
bone5 <- filter(data23, Bone == "Bone5")
bone6 <- filter(data23, Bone == "Bone6")
bone7 <- filter(data23, Bone == "Bone7")
bone8 <- filter(data23, Bone == "Bone8")
bone9 <- filter(data23, Bone == "Bone9")
bone10 <- filter(data23, Bone == "Bone10")
bone11 <- filter(data23, Bone == "Bone11")
bone12 <- filter(data23, Bone == "Bone12")
boxplot(bone1$Oxygen, bone2$Oxygen, bone3$Oxygen, bone4$Oxygen, bone5$Oxygen, bone6$Oxygen, bone7$Oxygen, bone8$Oxygen, bone9$Oxygen, bone10$Oxygen, bone11$Oxygen, bone12$Oxygen,
        main = "Oxygen Isotope Composition of Different T-rex Bones",
        xlab = "Bone",
        ylab = "Oxygen Isotope Level",
        names = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))
```

The boxplot above shows a comparison of the oxygen isotope levels in each of the 12 respective bones. It can be seen visually that there is a great deal of variability between the isotope levels of different bones. Initial observation indicates that the T-rex may be cold-blooded, due to the large fluctuation in temperatures. An F-test will provide further evidence of this observation.

```{r 23 anova, echo = FALSE}
#Perform an F-test for the different bone isotope levels
data23 <- (ex0523)
bonemodel <- lm(Oxygen ~ Bone, data23)
anova(bonemodel)
```

The ANOVA table above provides an analysis of the means and residuals of the oxygen levels of the 12 bones. From the ANOVA table, an extremely small *p*-value of **9.73e-07** is given, which provides convincing evidence that the means of oxygen isotope levels are different for the respective bones. This gives a strong indication that the T-rex was indeed cold-blooded.

These results may not infer causation because this was not a randomized experiment with bones randomly assigned to separate groups. Rather, the 12 bones were observed in their natural states. Additionally, these results may not be inferred to the larger population of Tyrannosaurus Rex specimens because these bones all came from a single T-rex specimen and there could be a number of confounding factors affecting oxygen isotope levels in this specimen such as environmental conditions, geographic location, physiological abnormalities, etc. The results may only be inferred to this single specimen. However, the results provide a good idea of what may be expected in other specimens if these tests were to be repeated.
